Simulated Observations of Young Protoplanetary Discs

### 2013-02-07 paper.tex
spell checked

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/continuum.eps}
 \caption{{\bf Left:}A 300$\,$GHz continuum image of the model, created from the average of ten LIME runs. {\bf Right:} the output of a single radiative transfer simulation, with artefacts due to finite gridding.}
 \label{averages}
\end{figure}

@@

OCS traces only the innermost 20$\,$AU of the disc and can be used to examine the central regions (figure \ref{sim_all}). As OCS is not seen in outflows (e.g. Stanke 07, Van der Tak 03) it can be used to view rotation of the central part of the disc without risk of contamination from emission from shocked material. The OCS line, which has the highest upper energy level among the selected transitions, traces the hottest and densest regions of the disc. As a result, the innermost spiral structure is best seen in OCS (26 - 25), as shown in figure \ref{sim_all}. The majority of the OCS line is visible in absorption against the bright continuum emission from the disc mid-plane. The exception to this is towards the centre of the disc where the hole in the hot, dense midplane means that there is little continuum to be absorbed.\newline

@@

The fractional abundace of C$^{17}$O is constant at 2$\times$10$^{-8}$ across the disc, meaning the C$^{17}$O line is the most accurate in reproducing the physical structure of the disc. Like H$_2$CO, C$^{17}$O shows emission in voids between spiral arms. The C$^{17}$O line is visible from inner edge of the disc to its outer edge. Like OCS, C$^{17}$O is not seen in outflows (e.g. yildiz et al. 2012) and so rotation seen in C$^{17}$O lines should only be due to rotation from the disc and not contamination from outflows.\newline

HCO$^+$ in unique in the molecules simulated in I2011 in that it traces only the outer regions of the disc, and so can be used to look at the extended velocity and physical structure. In figure \ref{sim_all} we can see that the HCO$^+$ line shows the outer edges of the spiral arms in absorption, but also shows emission from the diffuse gas further out in the disc. This allows the velocity structure of the disc measured out to larger radii. This allows constraints to be placed on the rotation of the disc out to larger raddi than other molecules.

@@

In order to make predictions on the observability of the synthetic brightness maps CASA (common astronomy software applications) was used to simulate their observtion with the completed ALMA in configuration 26 of 28, giving a beam size of between 0.02 and 0.03 arcseconds. As the weakest lines studied in the model are of the order of 0.1-0.2$\,$mJy$\,$beam$^{-1}$ at this resolution, sensitivities of the order of 0.02$\,$mJy$\,$beam$^{-1}$ are required. This requires around 6 hours of integration time. The exact sensitivities are give in table \ref{sigmas}.
\begin{table}
 \centering
 \begin{minipage}{80mm}
   \caption{Sensitivities of ALMA given 6 hours integration time with velocity resolution of $\sim$400$\,$ms$^{-1}$. As calculated by the ALMA on-line sensitivity calculator.}
   \label{sigmas}
   \begin{tabular}{c||c|c|c}
     \hline
     Species & Transition & Frequency & Sensitivity\\
     \hline
     C$^{17}$O & J=3$\rightarrow$2 & 337.06GHz & 0.0270$\,$mJy$\,$beam$^{-1}$ \\
     OCS & J=26$\rightarrow$25 & 316.15GHz & 0.0275$\,$mJy$\,$beam$^{-1}$ \\
     HCO$^+$ & J=3$\rightarrow$2 & 267.56GHz & 0.0176$\,$mJy$\,$beam$^{-1}$ \\
     H$_2$CO & 4$_{04}\rightarrow$3$_{03}$ & 290.62GHz & 0.0228$\,$mJy$\,$beam$^{-1}$ \\
     \hline
   \end{tabular}
 \end{minipage}
\end{table}
In order to clean the images which featured mainly absorption the sky model used was the output of the lime simulations with the continuum subtracted and then the image inverted. The simulated images were cleaned down to a 3$\sigma$ level with $\sigma$ given in table \ref{sigmas}.\newline

@@

Figure \ref{spectra} shows spectra from the lines simulated, one from the centre of the image and one on each side of the disc along the Y=0 axis. The most notable thing about these is, despite the fact that the spectra are equally spaced about the centre of the disc, they are not symmetric. Most noticeably in the C$^{17}$O the spectrum from the +X side of the disc is shows significantly more absorption than that of the -X side. In the OCS lines, the two sides are not moving with equal velocities and the line widths and shapes are markedly different. These can be atributed to the non-axissymmetric nature of the model.

@@

$\bullet$ CASA simulations of this model show that at a distance of 100$\,$pc extraction of kinematic and structural information from continuum and some molecular lines are possible in ALMA band 7.\newline 
$\bullet$ Our simulations show that many molecular species are predominantly seen in absorption against the hot mid-plane of self-gravitating protoplanetary discs, with emission only being seen where voids in the structure of the disc do not provide a bright continuum source to be absorbed.\newline 
$\bullet$ The quiescent nature of the envelope around such discs only obscures lines within $\pm\,$0.5 km$\,$s$^{-1}$ of the systematic velocity.\newline 
$\bullet$ The lines studied (OCS 26$\rightarrow$25, H$_2$CO 4$_{04}$$\rightarrow$3$_{03}$, C$^{17}$O 3$\rightarrow$2 and HCO$^+$ 3$\rightarrow$2) taken togther allow all regions of the disc to be sampled and the rotation of the disc to be constrained from the inner edge to the outer edge.\newline

>>>>

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/continuum.eps}
 \caption{{\bf Left:}A 300$\,$GHz continuum image of the model, created from the average of ten LIME runs. {\bf Right:} the output of a single radiative transfer simulation, with artifacts due to finite griding.}
 \label{averages}
\end{figure}

@@

OCS traces only the innermost 20$\,$AU of the disc and can be used to examine the central regions (figure \ref{sim_all}). As OCS is not seen in outflows (e.g. Stanke 07, Van der Tak 03) it can be used to view rotation of the central part of the disc without risk of contamination from emission from shocked material. The OCS line, which has the highest upper energy level among the selected transitions, traces the hottest and densest regions of the disc. As a result, the innermost spiral structure is best seen in OCS (26 - 25), as shown in figure \ref{sim_all}. The majority of the OCS line is visible in absorption against the bright continuum emission from the disc mid-plane. The exception to this is towards the centre of the disc where the hole in the hot, dense mid-plane means that there is little continuum to be absorbed.\newline

@@

The fractional abundance of C$^{17}$O is constant at 2$\times$10$^{-8}$ across the disc, meaning the C$^{17}$O line is the most accurate in reproducing the physical structure of the disc. Like H$_2$CO, C$^{17}$O shows emission in voids between spiral arms. The C$^{17}$O line is visible from inner edge of the disc to its outer edge. Like OCS, C$^{17}$O is not seen in outflows (e.g. Yildiz et al. 2012) and so rotation seen in C$^{17}$O lines should only be due to rotation from the disc and not contamination from outflows.\newline

HCO$^+$ in unique in the molecules simulated in I2011 in that it traces only the outer regions of the disc, and so can be used to look at the extended velocity and physical structure. In figure \ref{sim_all} we can see that the HCO$^+$ line shows the outer edges of the spiral arms in absorption, but also shows emission from the diffuse gas further out in the disc. This allows the velocity structure of the disc measured out to larger radii. This allows constraints to be placed on the rotation of the disc out to larger radii than other molecules.

@@

In order to make predictions on the observability of the synthetic brightness maps CASA (common astronomy software applications) was used to simulate their observation with the completed ALMA in configuration 26 of 28, giving a beam size of between 0.02 and 0.03 arcseconds. As the weakest lines studied in the model are of the order of 0.1-0.2$\,$mJy$\,$beam$^{-1}$ at this resolution, sensitivities of the order of 0.02$\,$mJy$\,$beam$^{-1}$ are required. This requires around 6 hours of integration time. The exact sensitivities are give in table \ref{sigmas}.
\begin{table}
 \centering
 \begin{minipage}{80mm}
   \caption{Sensitivities of ALMA given 6 hours integration time with velocity resolution of $\sim$400$\,$ms$^{-1}$. As calculated by the ALMA on-line sensitivity calculator.}
   \label{sigmas}
   \begin{tabular}{c||c|c|c}
     \hline
     Species & Transition & Frequency & Sensitivity\\
     \hline
     C$^{17}$O & J=3$\rightarrow$2 & 337.06GHz & 0.0270$\,$mJy$\,$beam$^{-1}$ \\
     OCS & J=26$\rightarrow$25 & 316.15GHz & 0.0275$\,$mJy$\,$beam$^{-1}$ \\
     HCO$^+$ & J=3$\rightarrow$2 & 267.56GHz & 0.0176$\,$mJy$\,$beam$^{-1}$ \\
     H$_2$CO & 4$_{04}\rightarrow$3$_{03}$ & 290.62GHz & 0.0228$\,$mJy$\,$beam$^{-1}$ \\
     \hline
   \end{tabular}
 \end{minipage}
\end{table}
In order to clean the images which featured mainly absorption the sky model used was the output of the lime simulations with the continuum subtracted and then the image inverted. The simulated images were cleaned down to a 3$\sigma$ level with $\sigma$ given in table \ref{sigmas}.\newline

@@

Figure \ref{spectra} shows spectra from the lines simulated, one from the centre of the image and one on each side of the disc along the Y=0 axis. The most notable thing about these is, despite the fact that the spectra are equally spaced about the centre of the disc, they are not symmetric. Most noticeably in the C$^{17}$O the spectrum from the +X side of the disc is shows significantly more absorption than that of the -X side. In the OCS lines, the two sides are not moving with equal velocities and the line widths and shapes are markedly different. These can be attributed to the non-axissymmetric nature of the model.

@@

$\bullet$ CASA simulations of this model show that at a distance of 100$\,$pc extraction of kinematic and structural information from continuum and some molecular lines are possible in ALMA band 7.\newline 
$\bullet$ Our simulations show that many molecular species are predominantly seen in absorption against the hot mid-plane of self-gravitating protoplanetary discs, with emission only being seen where voids in the structure of the disc do not provide a bright continuum source to be absorbed.\newline 
$\bullet$ The quiescent nature of the envelope around such discs only obscures lines within $\pm\,$0.5 km$\,$s$^{-1}$ of the systematic velocity.\newline 
$\bullet$ The lines studied (OCS 26$\rightarrow$25, H$_2$CO 4$_{04}$$\rightarrow$3$_{03}$, C$^{17}$O 3$\rightarrow$2 and HCO$^+$ 3$\rightarrow$2) taken together allow all regions of the disc to be sampled and the rotation of the disc to be constrained from the inner edge to the outer edge.\newline
